### 2022-03-29 manuscript/reports.Rmd

---
title: "Effectiveness of BNT162b2 booster doses in England: a observational study in OpenSAFELY-TPP: Materials"
output:
  html_document:
    keep_md: yes
    self_contained: yes
  word_document: default
  bookdown::html_document2:
    number_sections: no
    toc: no
  pdf_document:
    toc: no
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r setup, include=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')
source(here("lib", "functions", "utility.R"))

# where are the outputs (ie the inputs for this manuscript!) saved?
#output_dir_os <- here("output", "release-objects")
output_dir_os <- here("released_output", "release-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("manuscript", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

study_dates <-
  jsonlite::read_json(path=here("lib", "design", "study-dates.json")) %>%
  map(as.Date)

postbaselinecuts <- read_rds(here("lib", "design", "postbaselinecuts.rds"))

matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
matchsummary_treated <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 1L)
matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
matchflowchart <- read_csv(fs::path(output_dir_os, "matchflowchart.csv"))
matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
incidence <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0, yearsatrisk=yearsatrisk_1+yearsatrisk_0)

```

```{r, effects, echo=FALSE, message=FALSE, warning=FALSE}

  specify_decimal <- function(x, k, trim=FALSE) {

    fmtd <- format(round(x, k), nsmall = k)
    if (trim) {fmtd <- trimws(fmtd)}
    return(fmtd)
  }
  #
  # print_est1bracket <- function(x, b, round=1){
  #   paste0(specify_decimal(x, round), " (", specify_decimal(b, round), ")")
  # }
  #
  print_est2bracket <- function(x, b1, b2, round=1){
    paste0(specify_decimal(x, round), " (", specify_decimal(b1, round), ", ", specify_decimal(b2, round), ")")
  }

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidcc",
      "coviddeath",
      "noncoviddeath"
    ) %>% 
    set_names(., .)

# incidence <- 
#   bind_rows(
#     read_csv(fs::path(output_dir_os, "none", "incidence.csv")),
#     read_csv(fs::path(output_dir_os, "vax12_type", "incidence.csv")),
#     read_csv(fs::path(output_dir_os, "cev", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#     read_csv(fs::path(output_dir_os, "age65plus", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level))
#   ) %>%
#   filter(
#     outcome %in% outcomes
#   ) %>%
#   mutate(
#     subgroup_descr = fct_inorder(subgroup_descr),
#     subgroup_level = fct_inorder(subgroup_level),
#     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
#     #model_name = fct_inorder(str_wrap(model_name, 15)),
#   )

km <- 
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "km.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "km.csv")),
    read_csv(fs::path(output_dir_os, "cev", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(
    treatment = fct_inorder(treatment),
    treatment_descr = fct_inorder(treatment_descr),
    outcome = fct_inorder(outcome),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
    subgroup = fct_inorder(subgroup),
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_variable = fct_inorder(subgroup_variable),
    subgroup_variable_descr = fct_inorder(subgroup_variable_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    treated = (treated_descr=="Boosted")*1L
  ) %>%
  group_by(treatment, treated_descr, outcome, subgroup) %>%
  mutate(
    cml.risk = cumsum(n.risk),
    cml.event = cumsum(replace_na(n.event, 0)),
  )

groupvariables <- c("treatment", "treatment_descr", "outcome", "outcome_descr", "outcome_descr_wrap", "subgroup", "subgroup_descr", "subgroup_variable", "subgroup_variable_descr", "subgroup_level", "subgroup_level_descr")

km_incidence <-
  km %>%
  mutate(
    n.atrisk = n.risk,
    cml.atrisk = cumsum(replace_na(n.atrisk, 0)),
    cml.event = cumsum(replace_na(n.event, 0)),
    cml.censor = cumsum(replace_na(n.censor, 0)),
    rate = n.event/n.atrisk,
    cml.rate = cml.event / cml.atrisk,
    risk = 1 - surv,
  ) %>%
  select(
    treated,
    all_of(groupvariables),
    time, interval,
    surv, risk, n.atrisk, n.event, n.censor, rate, cml.atrisk, cml.event, cml.censor, cml.rate
  )

kmconstrast <- function(data, cuts=NULL){

  if(is.null(cuts)){cuts <- unique(data$time)}

  data %>%
    filter(time!=0) %>%
    mutate(
      period_start = cut(time, cuts, right=TRUE, label=cuts[-length(cuts)]),
      period_end = cut(time, cuts, right=TRUE, label=cuts[-1]),
      period = cut(time, cuts, right=TRUE, label=paste0(cuts[-length(cuts)]+1, " - ", cuts[-1]))
    ) %>%
    group_by(across(all_of(groupvariables))) %>%
    group_by(
      treated, period_start, period_end, period,
      .add = TRUE
    ) %>%
    summarise(
      interval = last(time) - first(time) + 1,
      time = last(time),
      cml.atrisk = last(cml.atrisk),
      cml.event = last(cml.event),
      cml.censor = last(cml.censor),
      cml.rate = last(cml.rate),
      persontime = sum(n.atrisk),
      n.atrisk = first(n.atrisk),
      n.event = sum(n.event, na.rm=TRUE),
      n.censor = sum(n.censor, na.rm=TRUE),
      surv = last(surv),
      risk = last(risk),
      rate = n.event/persontime,
    ) %>%
    pivot_wider(
      id_cols = c(groupvariables, "period_start", "period_end", "period",  "interval", "time"),
      names_from = treated,
      names_glue = "{.value}_{treated}",
      values_from = c(surv, risk, n.atrisk, n.event, n.censor, rate, cml.atrisk, cml.event, cml.censor, cml.rate)
    ) %>%
    mutate(
      kmrr = risk_1 / risk_0,
      kmrd = risk_1 - risk_0,
      irr = rate_1 / rate_0,
      ird = rate_1 - rate_0,
      cmlirr = cml.rate_1 / cml.rate_0,
      cmlird = cml.rate_1 - cml.rate_0
    ) %>%
    mutate(
      kmrd_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(kmrd),
      kmrd_inv_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(-kmrd),
    ) %>%
    mutate(
      cml.events = cml.event_0 + cml.event_1,
      cml.atrisk = cml.atrisk_0 + cml.atrisk_1
    )
}

kmtreated_daily <- kmconstrast(km_incidence)
kmtreated_cuts <- kmconstrast(km_incidence, postbaselinecuts)
kmtreated_overall <- kmconstrast(km_incidence, c(0,last(postbaselinecuts)))

km70 <- kmtreated_daily %>% filter(time==max(postbaselinecuts))

# 
# incidence <- 
#   bind_rows(
#     read_csv(fs::path(output_dir_os, "none", "incidence.csv")),
#     read_csv(fs::path(output_dir_os, "vax12_type", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#     #read_csv(fs::path(output_dir_os, "cev", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#     #read_csv(fs::path(output_dir_os, "age65plus", "incidence.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
#     #read_csv(fs::path(output_dir_os, "prior_covid_infection", "incidence.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
#   ) %>%
#   filter(
#     outcome %in% outcomes
#   ) %>%
#   mutate(
#     subgroup_descr = fct_inorder(subgroup_descr),
#     subgroup_level = fct_inorder(subgroup_level),
#     #subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
#   )

hazards <- 
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "prior_covid_infection", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(

    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
  )

overallhazards <- 
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "overalleffects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "overalleffects.csv")),
    read_csv(fs::path(output_dir_os, "cev", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "prior_covid_infection", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(

    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
  )

meta2hazards <- 
  # derive directly from hazards as meta2effects table needs to be re-run
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "prior_covid_infection", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(
    period = (term_right>28) + 1,
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
  ) %>%
  group_by(
    treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_variable, subgroup_descr, subgroup_level, subgroup_level_descr, model, model_descr, period, 
  ) %>%
  summarise(
    term_left = min(term_left),
    term_right = max(term_right),
    term_midpoint = (term_left+term_right)/2,
    estimate = weighted.mean(estimate, robust.se^-2),
    std.error = sqrt(1/sum(std.error^-2)),
    robust.se = sqrt(1/sum(robust.se^-2)),
    statistic = estimate/robust.se,
    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
    conf.low = estimate + qnorm(0.025)*robust.se,
    conf.high = estimate + qnorm(0.975)*robust.se,
    hr = exp(estimate),
    hr.ll = exp(conf.low),
    hr.ul = exp(conf.high),
    ve = 1-hr,
    ve.ll = 1-hr.ul,
    ve.ul = 1-hr.ll
  ) %>%
  ungroup() %>%
  mutate(
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
  )

# meta2hazards <- 
#   bind_rows(
#     read_csv(fs::path(output_dir_os, "none", "meta2effects.csv")),
#     read_csv(fs::path(output_dir_os, "vax12_type", "meta2effects.csv")),
#     read_csv(fs::path(output_dir_os, "cev", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#     read_csv(fs::path(output_dir_os, "age65plus", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#   ) %>%
#   filter(
#     outcome %in% outcomes
#   ) %>%
#   mutate(
#     subgroup_descr = fct_inorder(subgroup_descr),
#     subgroup_level = fct_inorder(subgroup_level),
#     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
#     #model_name = fct_inorder(str_wrap(model_name, 15)),
#     hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
#     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
#   )

hazards3 <- hazards %>% filter(model==3)

```

\newpage

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Table 0: Inclusion criteria

```{r flowchart, echo=FALSE, warning=FALSE, message=FALSE}

matchflowchart %>%
  transmute(
    treatment,
    stage,
    criteria = str_wrap(criteria, 40),
    n,
    n_exclude,
    pct_all
  ) %>%
  pivot_wider(
    id_cols = c(stage, criteria, treatment),
    names_from = treatment,
    values_from = c(n, n_exclude, pct_all)
  ) %>%
  gt() %>%
  cols_label(
    stage="",
    criteria = "",

    n_pfizer = "N",
    n_moderna = "N",

    n_exclude_pfizer = "N excluded",
    n_exclude_moderna = "N excluded",

    pct_all_pfizer = "% remaining",
    pct_all_moderna = "% remaining",
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("pfizer")
  ) %>%
   tab_spanner(
    label = "ChAdOx1-S",
    columns = ends_with("moderna")
  ) %>%
  # tab_spanner(
  #   label = "mRNA-1273",
  #   columns = ends_with("moderna")
  # ) %>%
  fmt_percent(
    columns = starts_with(c("pct_")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("n_")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "criteria")
  )
```

\newpage 
&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Table 1a: Participant characteristics for BNT162b2 boosting

Participant characteristics as on the day of recruitment into the treatment or control group.

```{r table1a, echo=FALSE, warning=FALSE, message=FALSE}
table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="pfizer")
table1_by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv")) %>% filter(treatment=="pfizer")

table1 <- 
  table1_body %>%
  #filter(outcome %in% outcomes) %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))

table1 %>%
  gt(groupname_col="treatment_descr") %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
     label=md("Characteristic")
  ) %>%
  cols_hide(c("variable", "var_label", "order"))
```

\newpage 
&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Table 1b: Participant characteristics for mRA-1273 boosting

Participant characteristics as on the day of recruitment into the treatment or control group.

```{r table1b, echo=FALSE, warning=FALSE, message=FALSE}
table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="moderna")
table1_by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv")) %>% filter(treatment=="moderna")

table1 <- 
  table1_body %>%
  #filter(outcome %in% outcomes) %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))

table1 %>%
  gt(groupname_col="treatment_descr") %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
     label=md("Characteristic")
  ) %>%
  cols_hide(c("variable", "var_label", "order"))
```

\newpage
&nbsp; &nbsp; &nbsp;

### Table 2a: Follow-up and outcomes for BNT162b2 boosting

Total follow-up and incidence rates for each outcome, by treatment group.

Counts in the underlying risk table has been rounded up to the nearest 6, so numbers may not add up exactly.

```{r table2a, echo=FALSE, warning=FALSE, message=FALSE}

format_ratio = function(numer,denom, width=7){
  paste0(
    replace_na(scales::comma_format(accuracy=1)(numer), "--"),
    " / ",
    str_pad(replace_na(scales::comma_format(accuracy=1)(denom),"--"), width=width, pad=" ")
  )
}

km70_table <- 
  kmtreated_overall %>%
  filter(
    #fup_period == "All",
    outcome %in% outcomes,
    treatment == "pfizer",
    #subgroup == "none",
  ) %>%
  ungroup() %>%
  mutate(
    n.atrisk = label_number(accuracy=1, big.mark = ",")(n.atrisk_1),
    q_0 = format_ratio(cml.event_0, cml.atrisk_0/365.25),
    q_1 = format_ratio(cml.event_1, cml.atrisk_1/365.25),
    risk_1=label_number(accuracy=0.01, scale=1000)(risk_1),
    risk_0=label_number(accuracy=0.01, scale=1000)(risk_0),
    kmrr = label_number(accuracy=0.001)(kmrr),
    kmrd = label_number(accuracy=0.01, scale=1000)(kmrd)
  ) %>%
  group_by(treatment, outcome, subgroup_variable_descr, subgroup_level_descr) %>%
  mutate(
    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, as.factor("")),
    subgroup_level_descr=if_else(subgroup_variable_descr!="Main analysis", subgroup_level_descr, as.factor(""))
  ) %>%
  group_by(treatment, outcome, subgroup_variable_descr) %>%
  mutate(
    subgroup_variable_descr=if_else(row_number()==1, subgroup_variable_descr, as.factor(""))
  ) %>% 
  ungroup() %>%
  select(
    outcome_descr,
    subgroup_variable_descr,
    subgroup_level_descr,
    n.atrisk,
    q_1,
    risk_1,
    q_0, 
    risk_0,
    kmrr,
    kmrd
  )

  gt(
    km70_table,
   groupname_col = c("outcome_descr"),
  ) %>%
  cols_label(
    subgroup_variable_descr = "Subgroup",
    subgroup_level_descr = "",
    #fup_period = "Days since booster",
    q_0 = "Events / person-years",
    q_1   = "Events / person-years",
    n.atrisk = "N per group",
    risk_0 = "Cumulative risk, per 1000 people",
    risk_1 = "Cumulative risk, per 1000 people",
    kmrr = "Risk ratio",
    kmrd = "Risk difference, per 1000 people"
  ) %>%
  tab_spanner(
    label = "Boosted",
    columns = ends_with("1")
  ) %>%
  tab_spanner(
    label = "Unboosted",
    columns = ends_with("0")
  ) %>%
  tab_spanner(
    label = "Boosted versus unboosted",
    columns = starts_with("km")
  ) %>%
  fmt_missing(
    everything(),
    missing_text=""
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  )

```
\newpage
&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Table 2b: Follow-up and outcomes for mRNA-1273 boosting

Total follow-up and incidence rates for each outcome, by treatment group.

Counts in the underlying risk table has been rounded up to the nearest 6, so numbers may not add up exactly.

```{r table2b, echo=FALSE, warning=FALSE, message=FALSE}

format_ratio = function(numer,denom, width=7){
  paste0(
    replace_na(scales::comma_format(accuracy=1)(numer), "--"),
    " / ",
    str_pad(replace_na(scales::comma_format(accuracy=1)(denom),"--"), width=width, pad=" ")
  )
}

km70_table <- 
  kmtreated_overall %>%
  filter(
    #fup_period == "All",
    outcome %in% outcomes,
    treatment == "moderna",
    #subgroup == "none",
  ) %>%
  ungroup() %>%
  mutate(
    n.atrisk = label_number(accuracy=1, big.mark = ",")(n.atrisk_1),
    q_0 = format_ratio(cml.event_0, cml.atrisk_0/365.25),
    q_1 = format_ratio(cml.event_1, cml.atrisk_1/365.25),
    risk_1=label_number(accuracy=0.01, scale=1000)(risk_1),
    risk_0=label_number(accuracy=0.01, scale=1000)(risk_0),
    kmrr = label_number(accuracy=0.001)(kmrr),
    kmrd = label_number(accuracy=0.01, scale=1000)(kmrd)
  ) %>%
  group_by(treatment, outcome, subgroup_variable_descr, subgroup_level_descr) %>%
  mutate(
    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, as.factor("")),
    subgroup_level_descr=if_else(subgroup_variable_descr!="Main analysis", subgroup_level_descr, as.factor(""))
  ) %>%
  group_by(treatment, outcome, subgroup_variable_descr) %>%
  mutate(
    subgroup_variable_descr=if_else(row_number()==1, subgroup_variable_descr, as.factor(""))
  ) %>% 
  ungroup() %>%
  select(
    #treatment_descr, 
    outcome_descr,
    subgroup_variable_descr,
    subgroup_level_descr,
    n.atrisk,
    q_1,
    risk_1,
    q_0, 
    risk_0,
    kmrr,
    kmrd
  )

  gt(
    km70_table,
   groupname_col = c("outcome_descr"),
  ) %>%
  cols_label(
    subgroup_variable_descr = "Subgroup",
    subgroup_level_descr = "",
    #fup_period = "Days since booster",
    q_0 = "Events / person-years",
    q_1   = "Events / person-years",
    n.atrisk = "N per group",
    risk_0 = "Cumulative risk, per 1000 people",
    risk_1 = "Cumulative risk, per 1000 people",
    kmrr = "Risk ratio",
    kmrd = "Risk difference, per 1000 people"
  ) %>%
  tab_spanner(
    label = "Boosted",
    columns = ends_with("1")
  ) %>%
  tab_spanner(
    label = "Unboosted",
    columns = ends_with("0")
  ) %>%
  tab_spanner(
    label = "Boosted versus unboosted",
    columns = starts_with("km")
  ) %>%
  fmt_missing(
    everything(),
    missing_text=""
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  )

```

\newpage
&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Table 3a: Estimated Pfizer booster effectiveness (5 period)

Main and subgroup analyses
```{r table_effectsa, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

bind_rows(
  hazards,
  overallhazards %>% mutate(term_left = 0, term_right=70)
) %>%
  filter(model==3, treatment=="pfizer") %>%
  mutate(
    period = fct_inorder(paste(term_left+1, "-", term_right)),
    ve_display = str_remove(ve_display, fixed("%"))
  ) %>%
  arrange(treatment_descr, outcome_descr, subgroup_level_descr, period) %>%
  group_by(outcome, treatment, subgroup_level_descr, model) %>%
  mutate(
    #subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr))),
    period = paste0(period, " days")
  ) %>%
  ungroup() %>%
  select(outcome_descr, subgroup_level_descr, period, ve_display) %>%
  pivot_wider(
    id_cols = c("outcome_descr", "subgroup_level_descr"),
    names_from = "period",
    #names_prefix ="period",
    values_from = "ve_display"
  ) %>%
  gt(
    groupname_col = c("outcome_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    subgroup_level_descr="",
    outcome_descr = "Outcome",
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  ) %>%
  tab_spanner(
    label = "Booster vaccine effectiveness (95% CI)",
    columns = ends_with("days")
  ) %>%
  fmt_missing(
    c("subgroup_level_descr"),
    missing_text=""
  )

```

\newpage
&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Table 3b: Estimated Moderna booster effectiveness (5 period)

Main and subgroup analyses
```{r table_effectsb, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

bind_rows(
  hazards,
  overallhazards %>% mutate(term_left = 0, term_right=70)
) %>%
  filter(model==3, treatment=="moderna") %>%
  mutate(
    period = fct_inorder(paste(term_left+1, "-", term_right)),
    ve_display = str_remove(ve_display, fixed("%"))
  ) %>%
  arrange(treatment_descr, outcome_descr, subgroup_level_descr, period) %>%
  group_by(outcome, treatment, subgroup_level_descr, model) %>%
  mutate(
    #subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr))),
    period = paste0(period, " days")
  ) %>%
  ungroup() %>%
  select(outcome_descr, subgroup_level_descr, period, ve_display) %>%
  pivot_wider(
    id_cols = c("outcome_descr", "subgroup_level_descr"),
    names_from = "period",
    #names_prefix ="period",
    values_from = "ve_display"
  ) %>%
  gt(
    groupname_col = c("outcome_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    subgroup_level_descr="",
    outcome_descr = "Outcome",
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  ) %>%
  tab_spanner(
    label = "Booster vaccine effectiveness (95% CI)",
    columns = ends_with("days")
  ) %>%
  fmt_missing(
    c("subgroup_level_descr"),
    missing_text=""
  )

```

\newpage

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Figure 1a: Pfizer booster vaccination uptake, treatment group eligibility and matching success

Cumulative booster vaccination over the duration of the study period, by matching eligibility and matching success. Separately by primary vaccination course.

```{r, vaxdatea, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='50%', results='asis', out.extra=''}
#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))

xmin <- min(matchcoverage$vax3_date )
xmax <- max(matchcoverage$vax3_date )+1
data_coverage <- matchcoverage  %>%
  mutate(
    status = factor(status, levels=c("ineligible", "unmatched", "matched")),
    status_descr = fct_relabel(status, str_to_title)
  ) %>%
  arrange(status) %>%
  mutate(
    status_descr = fct_inorder(status_descr)
  ) %>%
  filter(treatment=="pfizer") %>%
  mutate(
    # vax12_type = fct_case_when(
    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
    #   vax12_type == "az-az" ~ "ChAdOx1-S",
    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
    #   TRUE ~ NA_character_
    # ),
    course = fct_case_when(
      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2× BNT162b2  / 1× BNT162b2",
      vax12_type=="az-az" & treatment=="pfizer" ~ "2× ChAdOx1-S / 1× BNT162b2",
      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2× BNT162b2 / 1× mRNA-1273",
      vax12_type=="az-az" & treatment=="moderna" ~ "2× ChAdOx1-S / 1× mRNA-1273",
      TRUE ~ NA_character_
    )
  )

data_annotate <- 
  data_coverage %>%
  filter(vax12_type=="pfizer-pfizer") %>%
  group_by(status_descr, course) %>%
  summarise(
    xpos = first(lubridate::floor_date((vax3_date), "1 month")),
    ypos= max(cumuln)
  ) %>%
  ungroup() %>%
  arrange(desc(status_descr)) %>%
  mutate(
    ypos=(cumsum(ypos) + lag(cumsum(ypos), 1, 0))/2
  ) %>%
  mutate(
    course="2× ChAdOx1-S / 1× BNT162b2"
  )

ggplot(data_coverage) +
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=status_descr,
      fill=status_descr,
      colour=NULL
    ),
    position=position_stack(),
    alpha=0.5,
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
  geom_text(data=data_annotate, mapping=aes(x = xpos, y = (ypos), label = status_descr, colour=status_descr), hjust=0, show.legend = FALSE)+
  facet_wrap(vars(course))+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(data_coverage$vax3_date, "1 month")),
    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = scales::label_number(accuracy = 1, big.mark=","),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2", direction = -1)+
  scale_colour_brewer(type="qual", palette="Set2", direction = -1)+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "none"
  )+
  NULL
cat("  \n\n")
```

\newpage

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Figure 1b: Moderna booster vaccination uptake, treatment group eligibility and matching success

Cumulative booster vaccination over the duration of the study period, by matching eligibility and matching success. Separately by primary vaccination course.

```{r, vaxdateb, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='50%', results='asis', out.extra=''}
#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))

xmin <- min(matchcoverage$vax3_date )
xmax <- max(matchcoverage$vax3_date )+1
data_coverage <- matchcoverage  %>%
  mutate(
    status = factor(status, levels=c("ineligible", "unmatched", "matched")),
    status_descr = fct_relabel(status, str_to_title)
  ) %>%
  arrange(status) %>%
  mutate(
    status_descr = fct_inorder(status_descr)
  ) %>%
  filter(treatment=="moderna") %>%
  mutate(
    # vax12_type = fct_case_when(
    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
    #   vax12_type == "az-az" ~ "ChAdOx1-S",
    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
    #   TRUE ~ NA_character_
    # ),
    course = fct_case_when(
      #vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2× BNT162b2  / 1× BNT162b2",
      #vax12_type=="az-az" & treatment=="pfizer" ~ "2× ChAdOx1-S / 1× BNT162b2",
      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2× BNT162b2 / 1× mRNA-1273",
      vax12_type=="az-az" & treatment=="moderna" ~ "2× ChAdOx1-S / 1× mRNA-1273",
      TRUE ~ NA_character_
    )
  )

data_annotate <- 
  data_coverage %>%
  filter(vax12_type=="pfizer-pfizer") %>%
  group_by(status_descr, course) %>%
  summarise(
    xpos = first(lubridate::floor_date((vax3_date), "1 month")),
    ypos= max(cumuln)
  ) %>%
  ungroup() %>%
  arrange(desc(status_descr)) %>%
  mutate(
    ypos=(cumsum(ypos) + lag(cumsum(ypos), 1, 0))/2
  ) %>%
  mutate(
    course="2× ChAdOx1-S / 1× mRNA-1273"
  )

ggplot(data_coverage) +
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=status_descr,
      fill=status_descr,
      colour=NULL
    ),
    position=position_stack(),
    alpha=0.5,
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
  geom_text(data=data_annotate, mapping=aes(x = xpos, y = (ypos), label = status_descr, colour=status_descr), hjust=0, show.legend = FALSE)+
  facet_wrap(vars(course))+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(data_coverage$vax3_date, "1 month")),
    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = scales::label_number(accuracy = 1, big.mark=","),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2", direction = -1)+
  scale_colour_brewer(type="qual", palette="Set2", direction = -1)+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "none"
  )+
  NULL
cat("  \n\n")
```

\newpage

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

### Figure 2: Estimated booster effectiveness (5 period)

For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.

```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=9.5, out.width='90%', results='asis'}

formatpercent100 <- function(x,accuracy){
  formatx <- scales::label_percent(accuracy, suffix="")(x)

  if_else(
    formatx==scales::label_percent(accuracy, suffix="")(1),
    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
    formatx
  )
}

  ## hazard ratio

  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)

  overallhazards_plotdata <- 
    overallhazards %>%
    filter(
      model==3, 
      outcome %in% outcomes,
      #treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")
    ) %>%
    mutate(
      subgroup_level_descr=fct_relabel(subgroup_level_descr, ~str_wrap(., width=10))
    )

  plot_hazards <-
    hazards %>%
    filter(
      model==3,
      outcome %in% outcomes,
      #treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")
    ) %>%
    ungroup() %>%
    mutate(
      subgroup_level_descr_wrap=fct_relabel(subgroup_level_descr, ~str_wrap(.x, width=10))
    ) %>%
    ggplot()+
    #geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts+14)-7, xmax=max(postbaselinecuts+14)+7, fill="grey95", alpha=0.5, colour="transparent") +
    geom_point(aes(y=hr, x=term_midpoint, colour=treatment_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=treatment_descr), position = position_dodge(width = 1.8))+
    #geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts+14), colour=treatment_descr),  position = position_dodge(width = 1.8)) +
    #geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts+14), colour=treatment_descr),  position = position_dodge(width = 1.8)) +

    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(subgroup_level_descr_wrap), cols=vars(outcome_descr_wrap), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.05, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness (%)",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=c(postbaselinecuts), labels=c(postbaselinecuts), limits=c(min(postbaselinecuts), max(postbaselinecuts)), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=2))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_minimal()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement = "outside",
      strip.text.y.left = element_text(angle = 0),

      panel.spacing = unit(0.8, "lines"),

      plot.title = element_text(hjust = 0),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.caption = element_text(hjust = 0, face= "italic"),

      legend.position = "bottom",
      legend.direction = "horizontal"
     ) +
    NULL

  plot_hazards

```

\newpage

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

<!-- #### Figure 2b: Estimated booster effectiveness (2 period split at 28 days) -->

<!-- For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course. -->

<!-- ```{r plot_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'} -->

<!-- formatpercent100 <- function(x,accuracy){ -->
<!--   formatx <- scales::label_percent(accuracy, suffix="")(x) -->

<!--   if_else( -->
<!--     formatx==scales::label_percent(accuracy, suffix="")(1), -->
<!--     paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)), -->
<!--     formatx -->
<!--   ) -->
<!-- } -->
<!--  ## hazard ratio -->

<!--   y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2) -->

<!--   overallhazards_plotdata <-  -->
<!--     overallhazards %>% -->
<!--     filter( -->
<!--       model==3,  -->
<!--       outcome %in% outcomes, -->
<!--       treatment %in% c("pfizer"), -->
<!--       subgroup_variable %in% c("none", "vax12_type") -->
<!--     ) -->

<!--   postbaselinecuts28 <- c(0,28,70) -->

<!--   plot_hazards <- -->
<!--     meta28hazards %>% -->
<!--     filter( -->
<!--       model==3, -->
<!--       outcome %in% outcomes, -->
<!--       treatment %in% c("pfizer"), -->
<!--       subgroup_variable %in% c("none", "vax12_type") -->
<!--     ) %>% -->
<!--     ggplot()+ -->
<!--     geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
<!--     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
<!--     geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14)), colour="darkgrey") + -->
<!--     geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14)), colour="darkgrey") + -->
<!--     geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28+14)-7, xmax=max(postbaselinecuts28+14)+7, fill="grey95", alpha=0.5, colour="transparent") + -->
<!--     geom_hline(aes(yintercept=1), colour='grey')+ -->
<!--     facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+ -->
<!--     scale_y_log10( -->
<!--       breaks=y_breaks, -->
<!--       limits = c(0.01, 1.5), -->
<!--       oob = scales::oob_keep, -->
<!--       sec.axis = sec_axis( -->
<!--         ~(1-.), -->
<!--         name="Effectiveness (%)", -->
<!--         breaks = 1-(y_breaks), -->
<!--         labels = function(x){formatpercent100(x, 1)} -->
<!--       ) -->
<!--     )+ -->
<!--     scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, "Overall"), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+ -->
<!--     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+ -->
<!--     labs( -->
<!--       y="Hazard ratio", -->
<!--       x="Days since booster dose", -->
<!--       colour=NULL -->
<!--      ) + -->
<!--     theme_minimal()+ -->
<!--     theme( -->
<!--       panel.border = element_blank(), -->
<!--       axis.line.y = element_line(colour = "black"), -->
<!--       axis.line.y.right = element_blank(), -->

<!--       panel.grid.minor.x = element_blank(), -->
<!--       panel.grid.minor.y = element_blank(), -->
<!--       strip.background = element_blank(), -->
<!--       strip.placement = "outside", -->
<!--       strip.text.y.left = element_text(angle = 0), -->

<!--       panel.spacing = unit(0.8, "lines"), -->

<!--       plot.title = element_text(hjust = 0), -->
<!--       plot.title.position = "plot", -->
<!--       plot.caption.position = "plot", -->
<!--       plot.caption = element_text(hjust = 0, face= "italic"), -->

<!--       legend.position = "none" -->
<!--      ) + -->
<!--     NULL -->
<!--   plot_hazards -->

<!-- ``` -->

<!-- \newpage  -->

<!-- #### Figure 2c: Estimated booster effectiveness (2 period split at 42 days) -->

<!-- For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course. -->

<!-- ```{r plot_effects42, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'} -->

<!-- formatpercent100 <- function(x,accuracy){ -->
<!--   formatx <- scales::label_percent(accuracy, suffix="")(x) -->

<!--   if_else( -->
<!--     formatx==scales::label_percent(accuracy, suffix="")(1), -->
<!--     paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)), -->
<!--     formatx -->
<!--   ) -->
<!-- } -->
<!--  ## hazard ratio -->

<!--   y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2) -->

<!--   overallhazards_plotdata <-  -->
<!--     overallhazards %>% -->
<!--     filter( -->
<!--       model==3,  -->
<!--       outcome %in% outcomes, -->
<!--       treatment %in% c("pfizer"), -->
<!--       subgroup_variable %in% c("none", "vax12_type") -->
<!--     ) -->

<!--   postbaselinecuts42 <- c(0,42,70) -->

<!--   plot_hazards <- -->
<!--     meta42hazards %>% -->
<!--     filter( -->
<!--       model==3, -->
<!--       outcome %in% outcomes, -->
<!--       treatment %in% c("pfizer"), -->
<!--       subgroup_variable %in% c("none", "vax12_type") -->
<!--     ) %>% -->
<!--     ggplot()+ -->
<!--     geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
<!--     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
<!--     geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts42+14)), colour="darkgrey") + -->
<!--     geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts42+14)), colour="darkgrey") + -->
<!--     geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts42+14)-7, xmax=max(postbaselinecuts42+14)+7, fill="grey95", alpha=0.5, colour="transparent") + -->
<!--     geom_hline(aes(yintercept=1), colour='grey')+ -->
<!--     facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+ -->
<!--     scale_y_log10( -->
<!--       breaks=y_breaks, -->
<!--       limits = c(0.01, 1.5), -->
<!--       oob = scales::oob_keep, -->
<!--       sec.axis = sec_axis( -->
<!--         ~(1-.), -->
<!--         name="Effectiveness (%)", -->
<!--         breaks = 1-(y_breaks), -->
<!--         labels = function(x){formatpercent100(x, 1)} -->
<!--       ) -->
<!--     )+ -->
<!--     scale_x_continuous(breaks=c(postbaselinecuts42, max(postbaselinecuts42+14)), labels=c(postbaselinecuts42, "Overall"), limits=c(min(postbaselinecuts42), max(postbaselinecuts42+14)+14), expand = c(0, 0))+ -->
<!--     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+ -->
<!--     labs( -->
<!--       y="Hazard ratio", -->
<!--       x="Days since booster dose", -->
<!--       colour=NULL -->
<!--      ) + -->
<!--     theme_minimal()+ -->
<!--     theme( -->
<!--       panel.border = element_blank(), -->
<!--       axis.line.y = element_line(colour = "black"), -->
<!--       axis.line.y.right = element_blank(), -->

<!--       panel.grid.minor.x = element_blank(), -->
<!--       panel.grid.minor.y = element_blank(), -->
<!--       strip.background = element_blank(), -->
<!--       strip.placement = "outside", -->
<!--       strip.text.y.left = element_text(angle = 0), -->

<!--       panel.spacing = unit(0.8, "lines"), -->

<!--       plot.title = element_text(hjust = 0), -->
<!--       plot.title.position = "plot", -->
<!--       plot.caption.position = "plot", -->
<!--       plot.caption = element_text(hjust = 0, face= "italic"), -->

<!--       legend.position = "none" -->
<!--      ) + -->
<!--     NULL -->

<!--   plot_hazards -->

<!-- ``` -->

<!-- \newpage -->

<!-- ## Figure 3: Kaplan-Meier incidence curves -->

<!-- ```{r plot_km, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'} -->

<!-- plot_km <- km %>% -->
<!--   filter( -->
<!--     outcome %in% outcomes, -->
<!--     #treatment=="pfizer", -->
<!--     subgroup_variable %in% c("none", "vax12_type") -->
<!--   ) %>% -->
<!--     ggplot()+ -->
<!--     geom_step(aes(x=time, y=(1-surv)*1000, group=paste0(treated_descr,treatment_descr), alpha=treated_descr, colour=treatment_descr))+ -->
<!--     geom_rect(aes(xmin=time, xmax=time+1, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treatment_descr), alpha=0.1)+ -->
<!--     geom_hline(aes(yintercept=0), colour='black')+ -->
<!--     facet_grid(rows=vars(subgroup_level_descr), cols=vars(outcome_descr_wrap), switch="y", scales="free_y")+ -->
<!--     scale_x_continuous( -->
<!--       breaks = seq(0,7*max(postbaselinecuts),by=14), -->
<!--       expand = expansion(0) -->
<!--     )+ -->
<!--     scale_y_continuous( -->
<!--       expand = expansion(0), -->
<!--       #limits = c(0L,NA) -->
<!--     )+ -->
<!--     scale_colour_brewer(type="qual", palette="Set2", na.value="grey")+ -->
<!--     scale_fill_brewer(type="qual", palette="Set2", guide="none", na.value="grey")+ -->
<!--     labs( -->
<!--       x=NULL, -->
<!--       y="Cumulative incidence\nper 1000", -->
<!--       colour=NULL, -->
<!--       alpha=NULL -->
<!--     )+ -->
<!--     theme_minimal()+ -->
<!--     theme( -->
<!--       legend.position="bottom", -->
<!--       legend.justification = c(0,0), -->
<!--       legend.direction = "horizontal", -->
<!--       axis.text.x.top=element_text(hjust=0), -->

<!--       strip.placement = "outside", -->
<!--       strip.background = element_blank(), -->
<!--       strip.text.y.left = element_text(angle = 0), -->

<!--       panel.border = element_blank(), -->
<!--       axis.line.y = element_line(colour = "black"), -->
<!--       axis.line.x.top = element_line(colour = "black"), -->
<!--       panel.grid.minor.x = element_blank(), -->
<!--       panel.grid.minor.y = element_blank(), -->

<!--       panel.spacing = unit(0.8, "lines"), -->

<!--     )+ -->
<!--     NULL -->

<!-- plot_km -->

<!-- ``` -->